---
title: "HW3"
author: "Maurice"
output:
  html_document:
    df_print: paged
    toc: no
    toc_depth: '5'
  pdf_document:
    fig_height: 3
    fig_width: 4
    toc: no
    toc_depth: 5
header-includes: \usepackage{float} \usepackage{amsmath} \usepackage{amssymb}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

## Comment

Some of these exercises did not make any sense to me, and I have tried to interpret them as well as I could.

## Preamble

```{r}
# load libraries
require(odbc)
require(dbplyr)
require(tidyverse)
require(readxl)
require(knitr)
require(broom)

# connect to database
dbConn<- dbConnect(odbc(), 
                   driver = "SQL Server",
                   server = "ecoinfo2,6004",
                   database = "keesDB", uid = "TDM", pwd = "database")
```



## Exercise 1


*The COUNT of total number of individuals, BY Species and BY Year*
	

```{r}
query1 <- "
  SELECT 
    s.Species, 
    YEAR(m.SampDate) AS year, 
    COUNT(*) AS total
  FROM 
    tblFieldDemo AS m
  INNER JOIN 
    luSpecies AS s
  ON
    m.SpecCode = s.SpecCode
  GROUP BY 
    s.Species, 
    YEAR(m.SampDate);"

# execute query
rs = dbSendQuery(dbConn, query1)

# retrieve results and put into a variable
dbRows <- dbFetch(rs)

# Clear the result
dbClearResult(rs)

# put results in a formatted table
tbl1 <- spread(dbRows, year, total)
tbl1 <- kable(tbl1)
```


```{r echo = FALSE}
tbl1
```



\newpage
## Exercise 2

*The AVERAGE Monthly Temp_H BY Year*

Not sure what was meant, but I assume AVG montly temp, by year *and* month
	
```{r}
query2 <- "
  SELECT 
    YEAR(SampDate) AS yr, 
    MONTH(SampDate) AS mo, 
    AVG(Temp_H) AS avg_temp
  FROM 
    tblWeather
  GROUP BY 
    YEAR(SampDate), 
    MONTH(SampDate)
  ORDER BY 
    yr, 
    mo;"

# execute query
rs = dbSendQuery(dbConn, query2)

# retrieve results and put into a variable
dbRows <- dbFetch(rs)

# Clear the result
dbClearResult(rs)

# replace month names
dbRows <- dbRows %>%
  mutate(mo = case_when(mo == 1 ~ "Jan", 
                        mo == 2 ~ "Feb", 
                        mo == 5 ~ "May", 
                        mo == 7 ~ "Jul", 
                        mo == 8 ~ "Aug", 
                        mo == 10 ~ "Oct")) %>%
  rename(year = yr)

# build table
tbl2 <- spread(dbRows, mo, avg_temp)

# put columns in right order again
tbl2 <- tbl2 %>%
  select(year, Jan, Feb, May, Jul, Aug, Oct)

# put results in a formatted table
tbl2 <- kable(tbl2, digits = 3)
```


```{r echo = FALSE}
tbl2
```


\newpage
## Exercise 3


The AVERAGE Wt_g By Species and By Sex
	
```{r}
query3 <- "
  SELECT 
    Species, Sex, AVG(Wt_g) AS avg_weight
  FROM 
    tblFieldDemo AS f
  INNER JOIN 
    tblFieldMeas AS m
  ON 
    (f.PlotID = m.Plot AND 
    f.PlotArray = m.PlotArray AND 
    f.SampDate = m.SampDate AND 
    f.SpecCode = m.SpecCode AND 
    f.Rep = m.Rep)
  INNER JOIN 
    luSpecies AS s
  ON 
    f.SpecCode = s.SpecCode
  GROUP BY 
    Species, 
    Sex
  ORDER BY 
    Species, 
    Sex;"

# execute query
rs = dbSendQuery(dbConn, query3)

# retrieve results and put into a variable
dbRows <- dbFetch(rs)

# Clear the result
dbClearResult(rs)

# put results in a formatted table
tbl3 <- spread(dbRows, Sex, avg_weight)
tbl3 <- rename(tbl3, "Null"= `<NA>`)
tbl3 <- kable(tbl3, digits = 3)
```


```{r echo = FALSE}
tbl3
```


\newpage
## Exercise 4

*We are interested in creating a data.framethat can be used to fit an ANOVA (within R) for one particular species (SOLO). We want to know if SOLO abundance (total number of captures) is affected by Sex, Treatment (TRTMT), Block, Year and BasalArea_m2. Thus you will need to use multiple tables (JOINS) to obtain your required dataset. The format of the table should be in the following order (response variable ~ predictor variables):SOLO_Count, Sex, Treatment, Block, Year, BasalArea_m2*sdf

First correct the database, by putting in the right numbers fro BasalArea


```{r eval = FALSE}
# we need to update the basa area for each plot

# step 1 create new column in sql
  # added through SQL Server

# step 2 read in the spreadsheet since that is where the basal area lives
dat <- read_excel("SQLExample/SmallMammalData.xlsx", sheet = "FieldData")

# step 3: gget the unique BA values for each plot
ba <- dat %>%
  group_by(Plot) %>%
  summarize(BasalArea = mean(`Basal Area (m^2/ha)`, na.rm = TRUE))

# step 4: loop through the 12 rows of data, 
# and use an UPDATE query to change the values within the plot table
for(i in 1:nrow(ba)){
  dbSendQuery(dbConn, paste("UPDATE fcplots SET BasalArea_m2_ha = ", 
                            ba$BasalArea[i], 
                            "WHERE plotid = ", 
                            ba$Plot[i]))
}
```


Then start with the exercise

First, retrieve the data from the database

```{r}
query4 <- "
  SELECT 
    CodeDescription AS sex, 
    block, 
    trtmt, 
    YEAR(SampDate) AS yr,
    BasalArea_m2_ha AS BasalArea
  FROM 
    tblFieldDemo AS f
  INNER JOIN 
    fcplots
  ON 
    f.PlotID = fcplots.plotid
  INNER JOIN 
    luSex
  ON 
    f.Sex = luSex.Sex
  WHERE 
    SpecCode = 'SOLO';"

# execute query
rs = dbSendQuery(dbConn, query4)

# retrieve results and put into a variable
dbRows <- dbFetch(rs)

# Clear the result
dbClearResult(rs)

# calculate the count
rowCount <- nrow(dbRows)
dbRows$cnt <- rowCount

# put data in regression table
regData <- dbRows

# close database connection
dbDisconnect(dbConn)
```


Now do the regression

(I hope that this is the intended regression, because to me it doesn't make *any* sense)

\[ soloCount_i = sex_i + trtmt_i + block_i + year_i + basalArea_i + \epsilon_i \]

But because sex, trtmt, block and year are categorical, I need to transform them into dummies.


Find the dummies

```{r}
# prepare the data, make variables into dummy variables

# find which dummies I need
# for treatment
unique(regData$trtmt)

# for sex
unique(regData$sex)

# for block
unique(regData$block)

# for year
unique(regData$yr)
```

Now I need to create the dummies. I leave 

* sex: No Clue
* trtmt: SNAG
* block: 1
* yr: 2007

as defaults

Create the dummies

```{r}
# block: block1 is the default, then I need dummies for block2 and block3
regData <- regData %>%
  mutate(BLOCK2 = case_when(block == 2 ~ 1, TRUE ~ 0)) %>%
  mutate(BLOCK3 = case_when(block == 3 ~ 1, TRUE ~ 0))

# trtmt: SNAG is the default, then I need dummies for DWD, REM and CON
regData <- regData %>%
  mutate(DWD = case_when(trtmt == "DWD" ~ 1, TRUE ~ 0)) %>%
  mutate(REM = case_when(trtmt == "REM" ~ 1, TRUE ~ 0)) %>%
  mutate(CON = case_when(trtmt == "CON" ~ 1, TRUE ~ 0))

# sex: no clue is the default, then I need dummies for Female and Male
regData <- regData %>%
  mutate(FEMALE = case_when(sex == "Female" ~ 1, TRUE ~ 0)) %>%
  mutate(MALE = case_when(sex == "Male" ~ 1, TRUE ~ 0))

# year: 2007 is the default
regData <- regData %>%
  mutate(Y2008 = case_when(yr == 2008 ~1, TRUE ~ 0))
```




Do the regression

(which has become
\[soloCount_i = FEMALE_i + MALE_i + DWD_i + REM_i + CON_i +BLOCK2_i + BLOCK3_i + Y2008_i + basalArea_i + \epsilon_i \]
)

Where the default is block1, SNAG, No Clue, 2007 if all dummies are 0.


```{r}
# do regression
reg <- lm(cnt ~ FEMALE + MALE + DWD + REM + CON + BLOCK2 + BLOCK3 + Y2008 + BasalArea, regData)

# create anova data
anva <- aov(reg)

# put results in output tables
tblReg <- kable(tidy(reg), digits = 3)
tblAnova <- kable(tidy(anva), digits = 3)
```


```{r echo = FALSE}
tblReg
```



```{r echo = FALSE}
tblAnova
```


Or as was asked for, a summary table:

```{r}
summary(anva)
```


