---
title: "HW3"
author: "Maurice"
output:
  pdf_document:
    fig_height: 3
    fig_width: 4
    toc: no
    toc_depth: 5
  html_document:
    df_print: paged
    toc: no
    toc_depth: '5'
header-includes: \usepackage{float} \usepackage{amsmath} \usepackage{amssymb}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE)
```

## Comment

Some of these exercises did not make any sense to me, and I have tried to interpret them as well as I could.

Code can be found on my github page: www.github.com/mkees/ENR5050Coursework
And this document is also published on RPubs: www.rpubs.com/MKees/ENR505_HW3

## Preamble

```{r}
# load libraries
require(odbc)
require(dbplyr)
require(tidyverse)
require(readxl)
require(knitr)
require(broom)

# connect to database
dbConn<- dbConnect(odbc(), 
                   driver = "SQL Server",
                   server = "ecoinfo2,6004",
                   database = "keesDB", uid = "TDM", pwd = "database")
```



## Exercise 1


*The COUNT of total number of individuals, BY Species and BY Year*
	

```{r}
query1 <- "
  SELECT 
    s.Species, 
    YEAR(m.SampDate) AS year, 
    COUNT(*) AS total
  FROM 
    tblFieldDemo AS m
  INNER JOIN 
    luSpecies AS s
  ON
    m.SpecCode = s.SpecCode
  GROUP BY 
    s.Species, 
    YEAR(m.SampDate);"

# execute query
rs = dbSendQuery(dbConn, query1)

# retrieve results and put into a variable
dbRows <- dbFetch(rs)

# Clear the result
dbClearResult(rs)

# put results in a formatted table
tbl1 <- spread(dbRows, year, total)
tbl1 <- kable(tbl1)
```


```{r echo = FALSE}
tbl1
```



\newpage
## Exercise 2

*The AVERAGE Monthly Temp_H BY Year*

Not sure what was meant, but I assume AVG montly temp, by year *and* month
	
```{r}
query2 <- "
  SELECT 
    YEAR(SampDate) AS yr, 
    MONTH(SampDate) AS mo, 
    AVG(Temp_H) AS avg_temp
  FROM 
    tblWeather
  GROUP BY 
    YEAR(SampDate), 
    MONTH(SampDate)
  ORDER BY 
    yr, 
    mo;"

# execute query
rs = dbSendQuery(dbConn, query2)

# retrieve results and put into a variable
dbRows <- dbFetch(rs)

# Clear the result
dbClearResult(rs)

# replace month names
dbRows <- dbRows %>%
  mutate(mo = case_when(mo == 1 ~ "Jan", 
                        mo == 2 ~ "Feb", 
                        mo == 5 ~ "May", 
                        mo == 7 ~ "Jul", 
                        mo == 8 ~ "Aug", 
                        mo == 10 ~ "Oct")) %>%
  rename(year = yr)

# build table
tbl2 <- spread(dbRows, mo, avg_temp)

# put columns in right order again
tbl2 <- tbl2 %>%
  select(year, Jan, Feb, May, Jul, Aug, Oct)

# put results in a formatted table
tbl2 <- kable(tbl2, digits = 3)
```


```{r echo = FALSE}
tbl2
```


\newpage
## Exercise 3


The AVERAGE Wt_g By Species and By Sex
	
```{r}
query3 <- "
  SELECT 
    Species, Sex, AVG(Wt_g) AS avg_weight
  FROM 
    tblFieldDemo AS f
  INNER JOIN 
    tblFieldMeas AS m
  ON 
    (f.PlotID = m.Plot AND 
    f.PlotArray = m.PlotArray AND 
    f.SampDate = m.SampDate AND 
    f.SpecCode = m.SpecCode AND 
    f.Rep = m.Rep)
  INNER JOIN 
    luSpecies AS s
  ON 
    f.SpecCode = s.SpecCode
  GROUP BY 
    Species, 
    Sex
  ORDER BY 
    Species, 
    Sex;"

# execute query
rs = dbSendQuery(dbConn, query3)

# retrieve results and put into a variable
dbRows <- dbFetch(rs)

# Clear the result
dbClearResult(rs)

# put results in a formatted table
tbl3 <- spread(dbRows, Sex, avg_weight)
tbl3 <- rename(tbl3, "Null"= `<NA>`)
tbl3 <- kable(tbl3, digits = 3)
```


```{r echo = FALSE}
tbl3
```


\newpage
## Exercise 4

*We are interested in creating a data.framethat can be used to fit an ANOVA (within R) for one particular species (SOLO). We want to know if SOLO abundance (total number of captures) is affected by Sex, Treatment (TRTMT), Block, Year and BasalArea_m2. Thus you will need to use multiple tables (JOINS) to obtain your required dataset. The format of the table should be in the following order (response variable ~ predictor variables):SOLO_Count, Sex, Treatment, Block, Year, BasalArea_m2*


I do this exercise in several steps

1) correct the database
2) retrieve all data from database
3) build dependent variable *soloCount* in R
4) execute regression


### Correct the database

Here I correct the database, by putting in the correct values for *basalArea*, as was done in class. The code is not executed, but just shown.


```{r eval = FALSE}
# we need to update the basa area for each plot

# step 1 create new column in sql
  # added through SQL Server

# step 2 read in the spreadsheet since that is where the basal area lives
dat <- read_excel("SQLExample/SmallMammalData.xlsx", sheet = "FieldData")

# step 3: gget the unique BA values for each plot
ba <- dat %>%
  group_by(Plot) %>%
  summarize(BasalArea = mean(`Basal Area (m^2/ha)`, na.rm = TRUE))

# step 4: loop through the 12 rows of data, 
# and use an UPDATE query to change the values within the plot table
for(i in 1:nrow(ba)){
  dbSendQuery(dbConn, paste("UPDATE fcplots SET BasalArea_m2_ha = ", 
                            ba$BasalArea[i], 
                            "WHERE plotid = ", 
                            ba$Plot[i]))
}
```



### retrieve all data from database


```{r}
query4 <- "
  SELECT 
    CodeDescription AS sex, 
    block, 
    trtmt, 
    YEAR(SampDate) AS yr,
    BasalArea_m2_ha AS BasalArea
  FROM 
    tblFieldDemo AS f
  INNER JOIN 
    fcplots
  ON 
    f.PlotID = fcplots.plotid
  INNER JOIN 
    luSex
  ON 
    f.Sex = luSex.Sex
  WHERE 
    SpecCode = 'SOLO';"

# execute query
rs = dbSendQuery(dbConn, query4)

# retrieve results and put into a variable
regData <- dbFetch(rs)

# Clear the result
dbClearResult(rs)

# close database connection
dbDisconnect(dbConn)
```




### Build dependent variable

Now build the dependent variable *soloCount*


```{r}

# add in a dummy variable for number of observations over which I can sum
regData <- regData %>% 
  mutate(obs = 1)

# summarize data, and use "complete" to fill in missing values
summarizedRegData <- regData %>%
  complete(sex, block, trtmt, yr, fill = list(obs = 0)) %>%
  group_by(sex, block, trtmt, yr) %>%
  summarize(count = sum(obs))

# add in basal area, which was lost in the previous step
# (I'm sure it could have been done in one go, but I can't get it to work with "complete")

# create lookup table
luBasals <- regData %>%
  select(trtmt, BasalArea, block)

# keep only unique values
luBasals <- luBasals %>%
  unique() %>%
  arrange(trtmt, block)

# join basal area into the summarized data
summarizedRegData <- summarizedRegData %>% 
    left_join(luBasals, by = c("block" = "block", "trtmt" = "trtmt"))

```


### Execute regression


```{r}
# first put variables in correct format

# set variables as factors
summarizedRegData$sex <- as_factor(summarizedRegData$sex)
summarizedRegData$trtmt <- as_factor(summarizedRegData$trtmt)

# do regression
reg <- lm(count ~ sex + trtmt + block + yr + BasalArea, summarizedRegData)

# create anova data
anva <- aov(reg)

# put results in output tables
tblReg <- kable(tidy(reg), digits = 3)
tblAnova <- kable(tidy(anva), digits = 3)
```


The requested regression was
\[ soloCount_i = \alpha + sex_i + trtmt_i + block_i + year_i + basalArea_i + \epsilon_i \]

But because sex and trtmt are factors, R made them automatically into dummy variables. The regression has become
\[soloCount_i = \alpha + sexMale_i + sexNoClue_i + trtmtDWD_i + trtmtREM_i + trtmtSNAG_i + block + yr + BasalArea_i + \epsilon_i \]

Where the default is sexFemale and trtmtCON if all dummies are 0.


The regression results are as follows:

```{r echo = FALSE}
tblReg
```

And the anova table is:

```{r echo = FALSE}
tblAnova
```


Or as was asked for, a summary table:

```{r}
summary(anva)
```


